### 2012-05-07 paper/notes.tex
Added notes.tex

\documentclass[letterpaper, 11pt]{article}
\usepackage{graphicx}
\usepackage{natbib}
\usepackage[left=3cm,top=3cm,right=3cm]{geometry}

\renewcommand{\topfraction}{0.85}
\renewcommand{\textfraction}{0.1}
\newcommand{\stars}{\texttt{stars}}
\parindent=0cm

\title{Notes on Heirarchical Bayesian Inference}
\author{Brendon J. Brewer}

\begin{document}
\maketitle

\section{Introduction}
{\bf Question: Can we run a single inference and then do some kind of importance
sampling to get the inference with any proposed alternative hyperparameters?}

\section{No Hyperparameters}
For the StarField problem, we aim to infer the positions and fluxes of the
stars:
\begin{eqnarray}
\stars &=& \left\{x_i, y_i, f_i\right\}_{i=1}^N \\
&=& \left\{\theta_i\right\}_{i=1}^N
\end{eqnarray}

In software, we can sample from the following posterior PDF:
\begin{eqnarray}
p_0(\stars | D) \propto p_0(\stars) p(D|\stars)
\end{eqnarray}
where $p_0(\stars)$ is some kind of ignorance prior.

\section{Hyperparameters}
However, the posterior PDF we really want to sample is the following:
\begin{eqnarray}
p_1(\stars, \alpha | D) &\propto& p_1(\alpha)p_1(\stars|\alpha) p(D|\stars)\\
&=& p_1(\alpha)\left[\prod_{i=1}^N p_1(\theta_i|\alpha)\right] p(D|\stars)
\end{eqnarray}
where $\alpha$ denotes some hyperparameters. Can we get sampled catalogues from
$p_1$ by reusing the samples from $p_0$? Not by standard importance sampling -
dimensionality is too high.

\end{document}